---
title: "Combining maps"
author: "Louis Schroll"
date: "2024-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
options(knitr.table.NA = "")
```

```{r}
local_path <- "~/stage_M2/"
library(readxl)
library(tidyverse)
library(patchwork)
library(sf)
library(kableExtra)

load(file = paste0(local_path, "3.results/grid_spOccupancy_results.RData"))
load(file = paste0(local_path, "1.data/contour_golfe_du_lion.rdata"))
load(file = paste0(local_path, "1.data/wind_farm_boundaries.RData"))

indice_sensibilite_responsabilité <- read_excel(paste0(local_path,"6.source/indice_sensibilite_responsabilité.xlsx"), skip = 1)
score_df <- indice_sensibilite_responsabilité %>% 
  mutate(
    collision_score = Alt * (Man + Ptf + Noc) / 3,
    collision_summer = collision_score * CS_summer,
    collision_winter = collision_score * CS_winter,
    collision_summer_10    = round(collision_summer    / max(collision_summer, na.rm = T)    *10, 2),
    collision_winter_10    = round(collision_winter    / max(collision_winter, na.rm = T)    *10, 2)) %>% 
  mutate(
    displacement_score = (Dis * Spe),
    displacement_summer = displacement_score * CS_summer,
    displacement_winter = displacement_score * CS_winter,
    displacement_summer_10 = round(displacement_summer / max(displacement_summer, na.rm = T) *10, 2),
    displacement_winter_10 = round(displacement_winter / max(displacement_winter, na.rm = T) *10, 2)
  ) %>% 
  mutate(
    vul_furn = (Alt + Man + Ptf + Noc) * (Dis + Spe) / 8,
    vul_mean_summer = rowMeans(select(., displacement_summer_10, collision_summer_10)),
    vul_mean_winter = rowMeans(select(., displacement_winter_10, collision_winter_10)),
    vul_maxi_summer = pmax(collision_summer_10, displacement_summer_10),
    vul_maxi_winter = pmax(collision_winter_10, displacement_winter_10),
    vul_furn_summer = vul_furn * CS_summer,
    vul_furn_winter = vul_furn * CS_winter,
    vul_furn_summer = vul_furn_summer / max(vul_furn_summer, na.rm = T) * 10,
    vul_furn_winter = vul_furn_winter / max(vul_furn_winter, na.rm = T) * 10
  ) %>% 
  select(nom_fr, 
         collision_summer_10, collision_winter_10, 
         displacement_summer_10, displacement_winter_10,
         vul_mean_summer, vul_mean_winter,
         vul_maxi_summer, vul_maxi_winter,
         vul_furn_summer, vul_furn_winter) 


spOccupancy_res_grid <- spOccupancy_res_grid %>% 
  rename(mean_psi_labbe_HR = mean_psi_labbe,
         sd_psi_labbe_HR = sd_psi_labbe,
         mean_psi_oceanite_tempete_R = mean_psi_oceanite_tempete,
         sd_psi_oceanite_tempete_R = sd_psi_oceanite_tempete)

compute_vulnerability_gridcell <- function(results_grid, score_df, seasons = "summer", score_colname = "vul_furn_summer"){
  if (seasons == "summer"){
    species_names_end = "_R"
    extract_pattern = "(?<=mean_psi_)(.*?)(?=_R)"
  } else if (seasons == "winter"){
    species_names_end = "_HR"
    extract_pattern = "(?<=mean_psi_)(.*?)(?=_HR)"
  }
  mean_psi_df <- results_grid %>% 
    as_tibble() %>% 
    select(starts_with("mean_psi")) %>% 
    select(ends_with(species_names_end)) %>% 
    mutate(row_id = 1:nrow(.)) %>% 
    pivot_longer(-row_id, values_to = "mean_psi", names_to = "species_name") %>% 
    mutate(species_name = str_extract(string = species_name, pattern = extract_pattern)) 
  
  species_list <- unique(mean_psi_df$species_name)
  
  filtered_score <- score_df %>% filter(nom_fr %in% species_list)
  species_order <- filtered_score %>% pull(nom_fr)
  score_matrix <- filtered_score %>% pull(score_colname)
  
  mean_psi_matrix <- mean_psi_df %>% 
    pivot_wider(names_from = species_name, values_from = mean_psi) %>% 
    select(-row_id) %>% 
    relocate(all_of(species_order)) %>% 
    as.matrix()

  aggregated_value <- mean_psi_matrix %*% score_matrix

  return(aggregated_value)
}

plot_vulnerability_map <- function(grid_synthesis, 
                               fill_with,
                               legend_position = "bottom",
                               plot_title = "",
                               plot_title_size = 8,
                               legend_title_size = 9,
                               axis_text_size = 4,
                               plot_font = "Arial"){
  ggplot() +
    geom_sf(data = grid_synthesis, aes(fill = .data[[fill_with]]), color = NA, lwd = 0) +
    scale_fill_distiller(palette = "Spectral", 
                         breaks = c(min(grid_synthesis[[fill_with]]), max(grid_synthesis[[fill_with]])),
                         labels = c("Low", "High")) +
    geom_sf(data = contour_golfe, color = "black", fill = "antiquewhite") +
    geom_sf(data = wind_farm, fill = NA, col = "black") +
    #geom_sf(data = wind_farm, fill = NA, col = "black") +
    labs(title = plot_title) +
    theme_bw() +
    theme(
      #text = element_text(family = plot_font),
      legend.position = legend_position,
      legend.title = element_blank(),
      plot.title = element_text(hjust = 0.5, 
                                face = "bold", 
                                size = plot_title_size),
      plot.title.position = 'plot',
      axis.text = element_text(size = axis_text_size)
      ) +
    guides(
      fill = guide_colourbar(
        title = "Seabirds vulnerability",
        title.theme = element_text(
          face = "bold",
          size = legend_title_size,
         # family = plot_font,
          hjust = 0.5),
        title.position = 'top',
        ticks = FALSE,
        title.hjust = .5,
        barwidth = unit(6, 'lines'),
        barheight = unit(.4, 'lines'),
        label.theme = element_text(size = legend_title_size * 0.65)),
      colour = "none")
}
```


### Criteria used to compute risk :

Alt: estimated percentage of time flying at blade height

Man: Flight maneuverability

Ptf: Percentage of time spent flying

Noc: Nocturnal activity score

Dis: Sensibility score to disturbance by wind turbines, boats, helicopter

Spe: Habitat specialization score

CS: Conservation score (species status and importance of the area for the species)


```{r}
indice_sensibilite_responsabilité %>% 
  select(nom_francais, Alt, Man, Ptf, Noc, Dis, Spe, CS_summer, CS_winter) %>%
  kable(format = "markdown") %>%
  kable_styling()
```



### Vulnerabiliy

$$
Species \ vulnerability \ score = \frac{Alt + Man + Ptf + Noc}{4} \times \frac{Dis + Spe}{2} \times CS
$$

```{r}
score_df %>% 
  select(nom_fr, vul_furn_summer, vul_furn_winter) %>% #vul_mean_summer, vul_mean_winter, vul_maxi_summer, vul_maxi_winter,
  mutate(nom_fr = str_replace_all(string = nom_fr, pattern= "_", replacement = " ") %>% 
           str_to_sentence()) %>%
  kable(format = "markdown", 
        col.names = c("French name", "Summer", "Winter"),
        align = "c",
        caption = "Vulnerability score") %>%
  kable_styling() %>% 
  add_header_above(header = c(" " = 1, "Vulnerability" = 2))
```



#### Collision

$$
Collision \ risk \ score = \frac{Alt \times (Man + Ptf + Noc)}{3} \times CS
$$



```{r}
collision_df <- score_df %>% 
  select(nom_fr, collision_summer_10, collision_winter_10) %>% 
  mutate(nom_fr = str_replace_all(string = nom_fr, pattern= "_", replacement = " ") %>% 
           str_to_sentence(),
         collisison_average = rowMeans(select(., collision_summer_10, collision_winter_10)))

collision_df %>%
  kable(format = "markdown") %>%
  kable_styling()
```

#### Disturbance and displacements

$$
Displacement \ score = \frac{(Dis + Spe) \times CS}{10}
$$


```{r}
score_df %>% 
  select(nom_fr, displacement_summer_10, displacement_winter_10) %>% 
  mutate(nom_fr = str_replace_all(string = nom_fr, pattern= "_", replacement = " ") %>% 
           str_to_sentence(),
         collisison_average = rowMeans(select(., displacement_summer_10, displacement_winter_10))) %>%
  kable(format = "markdown") %>%
  kable_styling()
```


```{r, fig.height=5, fig.width=10}
grid_synthesis <- spOccupancy_res_grid %>%
  mutate(
    collision_summer = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "summer",
      score_colname = "collision_summer_10"
    ),
    collision_winter = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "winter",
      score_colname = "collision_winter_10"
    ),
    displacement_summer = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "summer",
      score_colname = "displacement_summer_10"
    ),
    displacement_winter = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "winter",
      score_colname = "displacement_winter_10"
    )) %>% 
  mutate(
    collision_all = rowMeans(select(as_tibble(.), 
                                    collision_summer, 
                                    collision_winter)),
    displacement_all = rowMeans(select(as_tibble(.), 
                                       displacement_summer, 
                                       displacement_winter))
  )



plot_collision_summer <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "collision_summer",
                                   plot_title = "Collision risk summer")

plot_collision_winter <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "collision_winter",
                                   plot_title = "Collision risk winter")

plot_collision_all <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "collision_all",
                                   plot_title = "Collision risk average")

plot_displacement_summer <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "displacement_summer",
                                   plot_title = "Displacement risk summer")

plot_displacement_winter <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "displacement_winter",
                                   plot_title = "Displacement risk winter")

plot_displacement_all <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "displacement_all",
                                   plot_title = "Displacement risk average")
```


```{r, fig.height=5, fig.width=5}
# (plot_collision_summer | plot_collision_winter | plot_collision_all) /
#   (plot_displacement_summer | plot_displacement_winter | plot_displacement_all)  + 
#   plot_layout(guides = "collect") +
#   plot_annotation(
#     title = "Seabirds vulnerability to wind farms",
#     #caption = 'Source: Migralion project & PELMED 2017-2021',
#     theme = theme(
#       plot.title = element_text(size = 10,
#                                 face = "bold",
#                                 hjust = 0.5),
#       plot.subtitle = element_text(hjust = 0.5),
#       legend.position = "bottom", 
#     ))
```



```{r, fig.height=10, fig.width=6}
grid_synthesis <- grid_synthesis %>%
  mutate(
    vulnerability_furness_summer = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "summer",
      score_colname = "vul_furn_summer"
    ),
    vulnerability_mean_summer = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "summer",
      score_colname = "vul_mean_summer"
    ),
    vulnerability_max_summer = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "summer",
      score_colname = "collision_summer_10"
    ),
    vulnerability_furness_winter = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "winter",
      score_colname = "vul_furn_winter"
    ),
    vulnerability_mean_winter = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "winter",
      score_colname = "vul_mean_winter"
    ),
    vulnerability_max_winter = compute_vulnerability_gridcell(
      results_grid = spOccupancy_res_grid,
      score_df = score_df,
      seasons = "winter",
      score_colname = "vul_maxi_winter"
    )
  ) %>% 
  mutate(
    vulnerability_furness_all = rowMeans(select(as_tibble(.), 
                                                 vulnerability_furness_summer, 
                                                 vulnerability_furness_winter)),
    vulnerability_mean_all = rowMeans(select(as_tibble(.), 
                                                 vulnerability_mean_summer, 
                                                 vulnerability_mean_winter)),
    vulnerability_max_all = rowMeans(select(as_tibble(.), 
                                                 vulnerability_max_summer, 
                                                 vulnerability_max_winter)))



plot_furness_summer <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "vulnerability_furness_summer",
                                   plot_title = "Furness formula summer")

plot_mean_summer <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                fill_with = "vulnerability_mean_summer",
                                plot_title = "Mean value summer")

plot_maxi_summer <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                               fill_with = "vulnerability_max_summer",
                               plot_title = "Max value summer")

plot_furness_winter <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "vulnerability_furness_winter",
                                   plot_title = "Furness formula winter")

plot_mean_winter <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                fill_with = "vulnerability_mean_winter",
                                plot_title = "Mean value winter")

plot_maxi_winter <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                               fill_with = "vulnerability_max_winter",
                               plot_title = "Max value winter")

plot_furness_all <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                   fill_with = "vulnerability_furness_all",
                                   plot_title = "Furness formula average")

plot_mean_all <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                                fill_with = "vulnerability_mean_all",
                                plot_title = "Mean value average")

plot_maxi_all <- plot_vulnerability_map(grid_synthesis = grid_synthesis,
                               fill_with = "vulnerability_max_all",
                               plot_title = "Max value average")
```


### Maps

```{r, fig.height=10, fig.width=9}

(plot_collision_summer | plot_collision_winter | plot_collision_all) /
  (plot_displacement_summer | plot_displacement_winter | plot_displacement_all) / 
  (plot_furness_summer | plot_furness_winter | plot_furness_all) /
  (plot_mean_summer | plot_mean_winter | plot_mean_all) /
  (plot_maxi_summer | plot_maxi_winter | plot_maxi_all) + 
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Seabirds vulnerability to wind farms",
    theme = theme(
      plot.title = element_text(size = 12,
                                face = "bold",
                                hjust = 0.5),
      legend.position = "bottom", 
    ))

```




```{r}
cities = data.frame(nom=c("Montpellier", "Perpignan", "Narbonne", "Marseille"), pays=rep("FR",4))
print(cities)

library(RJSONIO)
locateCountry = function(nameCity, codeCountry) {
  cleanCityName = gsub(' ', '%20', nameCity)
  url = paste(
    "http://nominatim.openstreetmap.org/search?city="
    , cleanCityName
    , "&countrycodes="
    , codeCountry
    , "&limit=9&format=json"
    , sep="")
  resOSM = fromJSON(url)
  if(length(resOSM) > 0) {
    return(c(resOSM[[1]]$lon, resOSM[[1]]$lat))
  } else return(rep(NA,2)) 
}

library(tidyverse)
coord = t(apply(cities, 1, function(aRow) locateCountry(aRow[1], aRow[2])))
cities_sf <- as_tibble(coord) %>% 
  rename(lat = V1, lon = V2) %>% 
  bind_cols(cities) %>% 
  st_as_sf(coords = c("lat", "lon"))

library("sf")
library("rnaturalearth")
library(ggspatial)

# https://r-spatial.org/r/2018/10/25/ggplot2-sf.html
world <- ne_countries(scale = "medium", returnclass = "sf")
st_crs(cities_sf) <- st_crs(world)
class(world)

area <- st_crop(
  x = world,
  xmin = 2.6,
  ymin = 42,
  xmax = 6,
  ymax = 44
)
```


```{r}
france <- st_crop(
  x = world,
  xmin = -4.5,
  ymin = 41,
  xmax = 8,
  ymax = 51
) %>% 
  filter(sovereignt != "United Kingdom")

rectangle <- st_bbox(grid2) %>% 
  as_tibble() %>% 
  mutate(name = c("xmin", "ymin", "xmax", "ymax")) %>% 
  pivot_wider(values_from = x, names_from = name) %>% 
  mutate(xmin = xmin - 0.15,
         xmax = xmax + 0.15,
         ymax = ymax + 0.15,
         ymin = ymin - 0.15)

map_france <- ggplot() + geom_sf(data = france, fill= "antiquewhite") +
  geom_rect(data = rectangle, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = NA, color = "red") +
  theme(panel.background = element_rect(fill = "aliceblue"),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.background = element_blank()) 
```


```{r}
city_names <- bind_cols(st_drop_geometry(cities_sf), 
                        st_coordinates(cities_sf)) %>% 
  mutate(Y = case_when(nom == "Narbonne" ~ Y - 0.06,
            TRUE ~ Y + 0.06))

grid2 <- st_transform(grid, st_crs(world))
lion_gulf_plot <- ggplot() + 
  geom_sf(data = grid2, lwd = 0.1, fill = NA) +
  geom_sf(data = area, fill= "antiquewhite") +
  geom_text(data=city_names, 
            aes(x=X, y=Y, label = nom),
            color = "black", 
            fontface = "bold", 
            check_overlap = T,
            size = 2) +
  # select an area
  # add scale (ggspatial package)
  annotation_scale(location = "br", width_hint = 0.3) +
  # add noth arrow (ggspatial package)
  annotation_north_arrow(location = "tr", which_north = "true", 
                        # pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) + 
  annotate(geom = "text", x = 4.5, y = 42.7, label = "Gulf of Lion", 
           fontface = "italic", color = "grey22", size = 6) +
  theme(panel.background = element_rect(fill = "aliceblue"),
        panel.grid = element_blank()) +
  geom_sf(data = cities_sf)
```


```{r}
library(patchwork)
lion_gulf_plot +
  inset_element(
    map_france,
    left = 0.1,
    bottom = 0.65,
    right = 0.4,
    top = 1,
    align_to = "full"
  )
```

